Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})

### 2014-05-20 manuscript.tex
methods JRI

%TO DO
%check citations in last paragraph for ter1
%reformat for AJB
%preprint somewhere
%finish fixing paragraph that jeff doesn't like
%check map details
%Code for SAS and command line stuff on Git
%Phenotyping and genotyping data on Git (it's on Figshare?)

@@

We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).  To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}. These analyses only included populations in which 8 or more individuals were sampled. To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}. BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}). We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3). Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis \citep{Pyhajarvi2013}.

@@

For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D). To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees. For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citealt{Chia2012}).

@@

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}). We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above. Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class are needed to detect the observed effect of the \emph{Hopscotch} on tillering index.

>>>>

%TO DO
%check citations in last paragraph for ter1
%reformat for AJB
%preprint somewhere
%finish fixing paragraph that jeff doesn't like
%check map details
%Code for SAS and command line stuff on Git
%Phenotyping and genotyping data on Git (it's on Figshare?) <- http://dx.doi.org/10.6084/m9.figshare.776926

@@

We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).  To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}. These analyses only included populations in which 8 or more individuals were sampled. To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}. BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}). We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3). Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.

@@

For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D). To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees. For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.

@@

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}). \jri{what is this url?} We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above. Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
